---
title: "R Plots for Lateral Transport in the Mediterranean"
output: html_notebook
---





Notebook for making pretty figures from the lateral transport study.

```{r}
library(tidyverse)
library(ggmap)
library(cmocean)
library(ggpubr)
```
Load data output from the batch analysis code.
```{r}

nemo_results_traj <- read_csv("../data/temp/nemo_results_traj.csv")
nemo_mean_ann_results <- read_csv("../data/temp/nemo_mean_ann_results.csv")
nemo_nov_may_results <- read_csv("../data/temp/nemo_nov_may_results.csv")
uk37_results_datatable <-read_csv("../data/temp/uk37_results_datatable.csv")
tex86_results_datatable <- read_csv("../data/temp/tex86_results_datatable.csv")

seas_matchup <- read.csv("../data/raw/Sediment_Locations_seas.csv")
```

Make the sinking speed into a factor and in the right order
```{r}
uk37_results_datatable <- uk37_results_datatable %>%
  transform(speed = gsub("^.{0,3}", "", speed)) %>%
  transform(speed = as.numeric(speed)) %>%
  filter(speed < 1000) %>%
  transform(speed = as.factor(speed))

tex86_results_datatable <- tex86_results_datatable %>%
  transform(speed = gsub("^.{0,3}", "", speed)) %>%
  transform(speed = as.numeric(speed)) %>%
  filter(speed < 1000) %>%
  transform(speed = as.factor(speed))
```

Match the keys for subbasin analysis
```{r}
seas_nemo_results <- left_join(nemo_mean_ann_results, select(seas_matchup, c(Location_key, Subbasin)), by = c("run_loc" = "Location_key"))
```

set the theme
```{r}
theme_set(theme_classic())
theme_update(axis.line = element_blank())
```


Subbasin analysis figure
```{r}
seas_check <- seas_nemo_results %>%
  filter(speed == "Sp_6") %>%
  ggplot(aes(
    x = site_lon,
    y = site_lat,
    color = Subbasin
  )) +
  geom_point() +
  annotation_map(map_data("world")) +
  coord_fixed() +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(name = NULL, limits = c(-6,35)) +
  scale_y_continuous(name = NULL, limits = c(30,46))

seas_check

ggsave("../results/figures/subbasin_map.pdf")
```

Plot the depth vs distance by subbasin
```{r}
depth_dist_seas <- seas_nemo_results %>%
  filter(speed == "Sp_6") %>%
  ggplot(aes(
    x = water_depth,
    y = mean_dist,
    color = Subbasin
  )) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point() +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(name = "Water depth (m)", expand = c(0, 0)) +
  scale_y_continuous(name = "Mean surface ocean to burial site distance (km)", expand = c(0, 0)) +
  ggtitle("Water depth and mean travel distance in Mediterranean subbasins, 6m/d sinking speed")

depth_dist_seas

ggsave("../results/figures/depth_dist_subbasin.pdf")
```

Depth vs transport bias by subbasin
```{r}
depth_offset_seas <- seas_nemo_results %>%
  filter(speed == "Sp_6") %>%
  ggplot(aes(
    x = water_depth,
    y = offset_ann_sst,
    color = Subbasin
  )) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point() +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(name = "Water depth (m)", expand = c(0, 0)) +
  scale_y_continuous(name = expression(paste(SST[end] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Transport bias (6 m" * d^{-1} * ")"))

depth_offset_seas

ggsave("../results/figures/depth_offset_subbasin.pdf")
```

Seasonal bias
```{r}
uk37_plots <- uk37_results_datatable %>%
  ggplot(aes(x = offset_nov_may_sst,
    y = proxy_offset_nov_may_site)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21, 22)) +
  stat_cor(cor.coef.name = c("R"), aes(fill = speed, label = ..r.label..)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = expression(paste("Nov to May ", SST[end] - SST[site], " (\u00b0C)"))) +
  scale_y_continuous(name = expression(paste("Nov to May ", SST[UK37] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled Nov-May transport bias at " * U[37]^{K} *" sites"))

uk37_plots

ggsave("../results/figures/uk37_seasonal_transport_bias_offsets.pdf")
```

Seasonal travel distance
```{r}
uk37_dist_plot <- uk37_results_datatable %>%
  ggplot(aes(x = mean_dist_nov_may,
    y = proxy_offset_nov_may_site)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21, 22)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = "Mean surface ocean to burial site distance (km)", expand = c(0, 0)) +
  scale_y_continuous(name = expression(paste("Nov to May ", SST[UK37] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled Nov-May transport distance and proxy offset for " * U[37]^{K} *" sites"))

uk37_dist_plot

ggsave("../results/figures/seasonal transport distance vs uk37 seasonal offset.pdf", uk37_dist_plot)
```

Modeled bias - UK37
```{r}
uk37_plots_ann <- uk37_results_datatable %>%
  ggplot(aes(x = offset_ann_sst,
    y = proxy_offset_site)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21, 22)) +
  labs(fill = "Sinking speed (m/d)") +
  stat_cor(cor.coef.name = c("R"), aes(fill = speed, label = ..r.label..)) +
  scale_x_continuous(name = expression(paste(SST[end] - SST[site], " (\u00b0C)"))) +
  scale_y_continuous(name = expression(paste(SST[UK37] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled annual transport bias at " * U[37]^{K} *" sites"))

uk37_plots_ann

ggsave("../results/figures/uk37_annual_transport_bias_offsets.pdf")
```

Travel distance - UK37
```{r}
uk37_dist_plot <- uk37_results_datatable %>%
  ggplot(aes(x = mean_dist_ann,
    y = proxy_offset_site)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21, 22)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = "Mean surface ocean to burial site distance (km)", expand = c(0, 0)) +
  scale_y_continuous(name = expression(paste(SST[UK37] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled mean annual transport distance and proxy offset for " * U[37]^{K} *" sites"))

uk37_dist_plot

ggsave("../results/figures/annual transport distance vs uk37 seasonal offset.pdf", uk37_dist_plot)
```

TEX86 transport bias
```{r}
tex86_offset_plot <- tex86_results_datatable %>%
  ggplot(aes(x = offset_ann_sst,
             y = proxy_offset_site_sst)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21, 22)) +
  stat_cor(cor.coef.name = c("R"), aes(fill = speed, label = ..r.label..)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = expression(paste(SST[end] - SST[site], " (\u00b0C)"))) +
  scale_y_continuous(name = expression(paste(SST[TEX86] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled transport bias at " * TEX[86] *" sites"))

tex86_offset_plot

ggsave("../results/figures/transport bias vs tex86 offset.pdf",tex86_offset_plot)
```

TEX86 travel distance
```{r}
tex86_dist_plot <- tex86_results_datatable %>%
  ggplot(aes(x = mean_dist,
             y = proxy_offset_site_sst)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21,22)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = "Mean surface ocean to burial site distance (km)", expand = c(0,0)) +
  scale_y_continuous(name = expression(paste(SST[TEX86] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled transport distance and proxy offset for "  * TEX[86] *" sites"))

tex86_dist_plot

ggsave("../results/figures/travel distance vs tex86 offset.pdf",tex86_dist_plot)
```

TEX86 transport bias for sites >1000m water depth
```{r}
tex86_offset_plot_deep <- tex86_results_datatable %>%
  filter(water_depth > 1000) %>%
  ggplot(aes(x = offset_ann_sst,
             y = proxy_offset_site_sst)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21,22)) +
  stat_cor(cor.coef.name = c("R"), aes(fill = speed, label = ..r.label..)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = expression(paste(SST[end] - SST[site], " (\u00b0C)"))) +
  scale_y_continuous(name = expression(paste(SST[TEX86] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled transport bias at " * TEX[86] *" sites deeper than 1000 m"))

tex86_offset_plot_deep

ggsave("../results/figures/modeled vs tex86 offset - only deep sites.pdf",tex86_offset_plot_deep)
```

TEX86 travel distance for sites >1000m water depth
```{r}
tex86_dist_plot_deep <- tex86_results_datatable %>%
  filter(water_depth > 1000) %>%
  ggplot(aes(x = mean_dist,
             y = proxy_offset_site_sst)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = speed, shape = speed), color = "black", stroke = 0.25) +
  scale_fill_cmocean("speed", discrete = TRUE) +
  scale_shape_manual(values = c(25,21,22,23,24,25,21,22)) +
  labs(fill = "Sinking speed (m/d)") +
  scale_x_continuous(name = "Mean surface ocean to burial site distance (km)", expand = c(0, 0)) +
  scale_y_continuous(name = expression(paste(SST[TEX86] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("Modeled transport distance and proxy offset for "  * TEX[86] *" sites deeper than 1000 m"))

tex86_dist_plot_deep

ggsave("../results/figures/travel distance vs tex86 offset - only deep sites.pdf",tex86_dist_plot_deep)
```

Transport bias at TEX86 sites, 6m/d sinking speed, color by water depth
```{r}
tex86_offset_sp6_plot <- tex86_results_datatable %>%
  filter(speed == 6) %>%
  ggplot(aes(x = offset_ann_sst,
             y = proxy_offset_site_sst,
             fill = water_depth)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = water_depth),
             shape = 21,
             color = "black",
             stroke = 0.25) +
  scale_fill_cmocean("deep", direction = -1) +
  guides(fill = guide_colorbar(title = "Water depth (m)")) +
  scale_x_continuous(name = expression(paste(SST[end] - SST[site], " (\u00b0C)"))) +
  scale_y_continuous(name = expression(paste(SST[TEX86] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("6 m/d transport bias at " * TEX[86] *" sites"))

tex86_offset_sp6_plot

ggsave("../results/figures/modeled vs tex86 offset - sp6.pdf",tex86_offset_sp6_plot)
```

Travl distance at TEX86 sites, 6m/d sinking speed, color by water depth
```{r}
tex86_dist_sp6_plot <- tex86_results_datatable %>%
  filter(speed == 6) %>%
  ggplot(aes(x = mean_dist,
             y = proxy_offset_site_sst,
             fill = water_depth)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point(aes(fill = water_depth),
             shape = 21,
             color = "black",
             stroke = 0.25) +
  scale_fill_cmocean("deep", direction = -1) +
  guides(fill = guide_colorbar(title = "Water depth (m)")) +
  scale_x_continuous(name = "Mean surface ocean to burial site distance (km)", expand = c(0, 0)) +
  scale_y_continuous(name = expression(paste(SST[TEX86] - SST[site], " (\u00b0C)"))) +
  ggtitle(expression("6 m/d transport distance and proxy offset for "  * TEX[86] *" sites"))

tex86_dist_sp6_plot

ggsave("../results/figures/travel dist vs tex86 offset - sp6.pdf",tex86_dist_sp6_plot)
```

Figure for the poster showing scatter plots of 6 m/d transport bias and proxy bias for both tex and uk
```{r} 
# Have to append the data to plot it all up with a column for proxy to make different color categories.
scatter_bias_plot <- bind_rows(
  'TEX86' = select(filter(
    tex86_results_datatable, speed ==6),
    c(Location_key,Latitude,Longitude,offset_ann_sst,proxy_offset_site_sst)),
  'UK37' = select(filter(
    uk37_results_datatable, speed == 6),
    c(Location_key,Latitude,Longitude,offset_ann_sst,proxy_offset_site_sst = proxy_offset_site)),
  .id = "Proxy") %>%
  ggplot(aes(x = offset_ann_sst,
             y = proxy_offset_site_sst,
             color = Proxy)) +
  geom_point(size = 4) +
  theme_bw(base_size = 15) +
  scale_color_manual(values = c('#bf5b17','#386cb0')) +
  scale_x_continuous(name = expression(paste("Transport bias (6 m/d) = ",SST[end] - SST[site], " (\u00b0C)"))) +
  scale_y_continuous(name = expression(paste("Proxy bias = ",SST[proxy] - SST[site], " (\u00b0C)")))

scatter_bias_plot

ggsave("../results/figures/transport bias vs proxy offset - sp6.pdf",scatter_bias_plot)

```
