---
title: "Transport bias from batch runs"
output: html_notebook
---

Takes in the output file from trajectoryfiledataextraction.py, the proxy database that matches up the model run locations with associated and nearby proxy data, and the information on coretop proxy data.

Outputs files that are then used for plotting scripts and summary tables.


```{r}
library(tidyverse)

```



Load:
    the database that lists all sites and their associated proxies
    the file with extracted trajectory endpoint information
    files with the core-top data for each proxy

```{r}
proxy_db <- read_csv('../data/raw/Sediment_Locations.csv')
sst_data <- read_csv('../data/processed/extracted_trajectory_info.csv')


uk37_coretop <- read_csv('../data/raw/uk37_table.csv')
tex86_coretop <- read_csv('../data/raw/tex_table.csv')
```





Grab the site IDs. First need to create a table with just the locations that are run locations, then match on that.
```{r}
run_locs <- proxy_db %>%
  filter(proxy_db$Location_key == proxy_db$run_loc) %>%
  select(c("run_loc","Latitude","Longitude"))
  
sst_data <- left_join(sst_data, run_locs, by = c("site_lat" = "Latitude", "site_lon" = "Longitude"))

```




Make sure the deep data is numeric.
```{r}
nemo_ssts <- sst_data %>%
  transform(endpt_sst = as.numeric(endpt_sst),
            deep_sst = as.numeric(deep_sst))
```



Calculate the proxy-based SSTs for comparison with the site and trajectory endpoint information.

UK37: T = (UK37 - 0.039)/0.034 (Prahl et al. 1988)

TEX86: T = 38.6+68.4*TEX86 (Kim et al. 2010)

```{r}
uk37_coretop <- uk37_coretop %>%
  mutate(uk37_sst = (UK37 - 0.039)/0.034)
tex86_coretop <- tex86_coretop %>%
  mutate(tex86_sst = 38.6 + 68.4 * log10(TEX86))
```






Get proxy-based subsets for matching up with the NEMO data
```{r}
uk_sites <- filter(proxy_db, (!is.na(proxy_db$UK37_ID)))
tex_sites <- filter(proxy_db,(!is.na(proxy_db$TEX86_ID)))
```





Now get the annual and seasonal means .
```{r}
nemo_means <- nemo_ssts %>%
  group_by(run_loc,site_lat,site_lon,speed,water_depth) %>%
  summarise(ann_sst = mean(endpt_sst),
            ann_deep_sst = mean(deep_sst),
            mean_dist = mean(dist_final),
            mean_deep_dist = mean(dist_150m))

nemo_nov_may <- nemo_ssts %>%
  filter(months_end %in% c(11,12,1,2,3,4,5)) %>%
  group_by(run_loc,site_lat,site_lon,speed,water_depth) %>%
  summarise(nov_may_sst = mean(endpt_sst),
            nov_may_deep = mean(deep_sst),
            mean_dist = mean(dist_final),
            mean_deep_dist = mean(dist_150m))

```







Now calculate the offset, (end - site), where site is represented by the 1000 m/d sinking speed

Filter the data to make separate tables for the site (1000 m/d) data, then add column for site data to the data table, then calculate the offset.
```{r}
site_mean_ann <- nemo_means %>%
  filter(speed == "Sp_1000") %>%
  select(run_loc, ann_sst, ann_deep_sst) %>%
  rename(site_sst = ann_sst,
         site_deep_sst = ann_deep_sst,
         site_sp = speed)

site_nov_may <- nemo_nov_may %>%
  filter(speed == "Sp_1000") %>%
  select(run_loc, nov_may_sst, nov_may_deep) %>%
  rename(site_nov_may_sst = nov_may_sst,
         site_deep_nov_may_sst = nov_may_deep,
         site_sp = speed)

```



```{r}
nemo_means_all <- left_join(nemo_means, site_mean_ann, by = c("run_loc","site_lat", "site_lon"))

nemo_nov_may_all <- left_join(nemo_nov_may, site_nov_may, by = c("run_loc","site_lat", "site_lon"))
```



```{r}
nemo_mean_offset <- nemo_means_all %>%
  mutate(offset_ann_sst = ann_sst - site_sst,
         offset_ann_deep = ann_deep_sst - site_deep_sst)

nemo_nov_may_offset <- nemo_nov_may_all %>%
  mutate(offset_nov_may_sst = nov_may_sst - site_nov_may_sst,
         offset_nov_may_deep = nov_may_deep - site_deep_nov_may_sst)
```





Match the site IDs, run IDs, and get proxy-based subsets
```{r}
uk_ssts_ann_nemo <- left_join(uk_sites, nemo_mean_offset, by = "run_loc")
tex_ssts_ann_nemo <- left_join(tex_sites, nemo_mean_offset, by = "run_loc")
uk_ssts_nov_may_nemo <- left_join(uk_sites, nemo_nov_may_offset, by = "run_loc")

uk37_ann_nemo_results <- left_join(uk_ssts_ann_nemo, uk37_coretop, by = c("UK37_ID"))
tex86_ann_nemo_results <- left_join(tex_ssts_ann_nemo, tex86_coretop, by = c("TEX86_ID"))
uk37_nov_may_nemo_results <- left_join(uk_ssts_nov_may_nemo, uk37_coretop, by = c("UK37_ID"))
```









Calculate the proxy offsets
```{r}
uk37_ann_nemo_results <- uk37_ann_nemo_results %>%
  mutate(proxy_offset = uk37_sst - ann_sst,
         proxy_offset_site = uk37_sst - site_sst)

tex86_ann_nemo_results <- tex86_ann_nemo_results %>%
  mutate(proxy_offset_sst = tex86_sst - ann_sst,
         proxy_offset_deep = tex86_sst - ann_deep_sst,
         proxy_offset_site_sst = tex86_sst - site_sst,
         proxy_offset_site_deep = tex86_sst - site_deep_sst)

uk37_nov_may_nemo_results <- uk37_nov_may_nemo_results %>%
  mutate(proxy_offset_nov_may = uk37_sst - nov_may_sst,
         proxy_offset_nov_may_site = uk37_sst - site_nov_may_sst)
```







Get things ready for plotting in python


```{r}
uk37_ann_nemo_wider <- select(uk37_ann_nemo_results, c("Location_key", "Latitude", "Longitude","water_depth", "speed","site_sst","uk37_sst","proxy_offset_site","ann_sst","mean_dist", "offset_ann_sst","proxy_offset")) %>%
  pivot_wider(names_from = c(speed), values_from = c("ann_sst":"proxy_offset"))


tex86_ann_nemo_wider <- select(tex86_ann_nemo_results, c("Location_key", "Latitude", "Longitude","water_depth", "speed","ann_sst":"mean_deep_dist","site_sst":"offset_ann_deep","tex86_sst":"proxy_offset_site_deep")) %>%
  pivot_wider(names_from = c(speed), values_from = c("ann_sst":"mean_deep_dist","offset_ann_sst":"offset_ann_deep","proxy_offset_sst","proxy_offset_deep"))

uk37_nov_may_nemo_wider <- select(uk37_nov_may_nemo_results, c("Location_key", "Latitude","water_depth", "Longitude","uk37_sst","site_nov_may_sst","proxy_offset_nov_may_site", "speed","nov_may_sst","mean_dist","offset_nov_may_sst","proxy_offset_nov_may")) %>%
  pivot_wider(names_from = c(speed), values_from = c("nov_may_sst":"proxy_offset_nov_may"))

```




Export to csv
```{r}
write_csv(uk37_ann_nemo_wider,"../data/temp/uk37_ann_nemo_to_plot.csv")
write_csv(tex86_ann_nemo_wider,"../data/temp/tex86_ann_nemo_to_plot.csv")

write_csv(uk37_nov_may_nemo_wider,"../data/temp/uk37_nov_may_nemo_to_plot.csv")
```






Get all the uk37 stuff into one spot for plotting
```{r}

uk37_all_nemo <- left_join(
  select(uk37_ann_nemo_results,c(
    Location_key, Latitude, Longitude, speed, ann_sst, mean_dist, site_sst, offset_ann_sst, wdepth, uk37_sst, proxy_offset, proxy_offset_site
  )), 
  select(uk37_nov_may_nemo_results,c(
    Location_key, speed, mean_dist, site_nov_may_sst, nov_may_sst, offset_nov_may_sst, proxy_offset_nov_may, proxy_offset_nov_may_site
  )), 
  by = c("Location_key","speed"))

uk37_all_nemo <- uk37_all_nemo %>%
  rename(mean_dist_ann = mean_dist.x,
         mean_dist_nov_may = mean_dist.y) 
  
```



Pivot wider on the nemo sst tables for easy plotting in the python notebook
```{r}
nemo_mean_offset_wider <- nemo_mean_offset %>%
  rename("Latitude" = "site_lat", "Longitude" = "site_lon") %>%
  select(c(run_loc, Latitude, Longitude, water_depth, speed, mean_dist:offset_ann_deep)) %>%
  pivot_wider(names_from = c(speed),
              values_from = c(mean_dist, mean_deep_dist, offset_ann_sst, offset_ann_deep))

nemo_nov_may_offset_wider <- nemo_nov_may_offset %>%
  rename("Latitude" = "site_lat", "Longitude" = "site_lon") %>%
  select(c(run_loc, Latitude, Longitude, water_depth, speed, mean_dist:offset_nov_may_deep)) %>%
  pivot_wider(names_from = c(speed),
              values_from = c(mean_dist, mean_deep_dist, offset_nov_may_sst, offset_nov_may_deep))
```





Write everything to csv for plotting in a different script
```{r}
write_csv(uk37_all_nemo,"../data/temp/uk37_results_datatable.csv")
write_csv(tex86_ann_nemo_results, "../data/temp/tex86_results_datatable.csv")

write_csv(nemo_mean_offset_wider,"../data/temp/nemo_mean_ann_results_wider.csv")
write_csv(nemo_nov_may_offset_wider,"../data/temp/nemo_nov_may_results_wider.csv")

write_csv(nemo_mean_offset,"../data/temp/nemo_mean_ann_results.csv")
write_csv(nemo_nov_may_offset,"../data/temp/nemo_nov_may_results.csv")

write_csv(nemo_ssts,"../data/temp/nemo_results_traj.csv")
```






