---
title: "Directionality analysis"
output: html_notebook
---

This notebook will calculate the bearing of transported particles and use a wind rose plot to visualize the bearing and distance of transported particles.

```{r}
library(tidyverse)

library(geosphere, include.only = c("bearingRhumb"))
```

Load the raw data to calculate bearing from each site-endpoint pair.
```{r}
traj_data <- read_csv("~/Documents/CodeRepo/lat-trans-med/data/processed/extracted_trajectory_info.csv", 
    col_types = cols(`150m_lat` = col_double(), 
        `150m_lon` = col_double()))
```

Calculate the bearing from the site to the endpoint for both surface and subsurface.
```{r}
traj_data <- na.omit(traj_data) %>%
  rename(deep_lon = "150m_lon",
         deep_lat = "150m_lat") %>%
  rowwise %>%
  mutate(
    bearing_surface = bearingRhumb(c(site_lon, site_lat), c(end_lon, end_lat)),
    bearing_subsurface = bearingRhumb(c(site_lon, site_lat), c(deep_lon,deep_lat))
  )
```

Calculate the difference in transport distance, distance between endpoint and deep endpoint, and difference in bearing angle.
```{r}
traj_data <- na.omit(traj_data) %>%
  rowwise %>%
  mutate(travel_dist_diff = dist_final - dist_150m) %>%
  mutate(dist_end_deep = geosphere::distm(c(deep_lat,deep_lon),c(end_lat,end_lon))) %>%
  mutate(bearing_diff = if(abs(bearing_surface - bearing_subsurface) > 180) (
      360 - abs(bearing_surface - bearing_subsurface)
    ) else (
      abs(bearing_surface - bearing_subsurface)
    )
  )
```

plot of bearings
```{r}
bearings_plot <- na.omit(traj_data) %>%
  filter(speed == "Sp_6") %>%
  ggplot(aes(x = bearing_surface,
      y = bearing_subsurface,
      color = site_lon)
      ) +
  geom_point()

bearings_plot
```

summarize by site and speed and take the travel distance weighted mean and the mean difference in bearing
```{r}
traj_data <- ungroup(na.omit(traj_data)) %>%
  mutate(circ_bearing_surface = circular::circular(bearing_surface, units = "degrees"),
         circ_bearing_subsurface = circular::circular(bearing_subsurface, units = "degrees"))
bearings_summary <- traj_data %>%
  group_by(site_lat,site_lon,speed,water_depth) %>%
  summarise(
    mean_dist_final = mean(dist_final),
    mean_dist_deep = mean(dist_150m),
    count_traj = n(),
    mean_bearing = circular::weighted.mean.circular(circ_bearing_surface,
                                                    dist_final/(mean_dist_final*count_traj)),
    mean_bearing_deep = circular::weighted.mean.circular(circ_bearing_subsurface,
                                                    dist_150m/(mean_dist_deep*count_traj))
  ) %>%
  mutate(mean_bearing_diff = if(abs(mean_bearing - mean_bearing_deep) > 180) (
    360 - abs(mean_bearing - mean_bearing_deep)
    ) else (
      abs(mean_bearing - mean_bearing_deep)))
```

plot it again?
```{r}
bearing_summary_plot <- bearings_summary %>%
  filter(speed == "Sp_6") %>%
  ggplot(aes(x = site_lon,
             y = site_lat,
             color = mean_bearing_diff)) +
  geom_point() +
  annotation_map(map_data("world")) +
  coord_fixed()

bearing_summary_plot
ggsave("../results/figures/bearing_surface_subsurface.pdf")
```

```{r}
bearing_hist <- ggplot(data = traj_data, aes(x = bearing_diff)) + geom_histogram()

bearing_hist
```

